Adolescent well-being in the digital age

### 2022-02-01 models.Rmd
Use .Rmd
Estimate models in script

# Model fitting {.unnumbered}

```{r setup}
#| cache: false
# Packages
library(scales)
library(ggtext)
library(lubridate)
library(kableExtra)
library(ggstance)
library(mgcv)
library(brms)
library(cmdstanr)
library(ellipse)
library(ggh4x)
library(tidybayes)
library(ggdist)
library(posterior)
library(parameters)
library(emmeans)
library(broom)
library(tidyverse)

# MCMC settings (ensure at most 12 cores)
ncores <- min(parallel::detectCores(logical = FALSE), 12)
nchains <- 4
options(brms.backend = "cmdstanr")
hmc <- list(
  chains = nchains,
  cores = nchains,
  threads = ncores %/% nchains,
  iter = 2500,
  warmup = 1000,
  refresh = 50,
  adapt_delta = .90,
  max_treedepth = 10
)

# Create directory for intermediate files
dir.create("models", FALSE)

# Load data sets
itu <- read_rds("data/itu.rds")
gwp <- read_rds("data/gwp.rds")
gbd <- read_rds("data/gbd.rds")
```

@@

We first prepare the variables for modelling:

-   Create lagged and centered predictors
-   Transform edge values of internet variable to fit beta
    -   2 values were changed (0 in Timor-Leste in 2002, and 1 in UAE 2020)
-   Join outcome and predictor tables
-   Center year on 2010 (roughly in the middle)
-   Ensure that age and sex are treated with contrast codes

```{r model1-prepare-data}
# Create lagged predictors
itu <- itu %>% 
  arrange(region, country, year) %>%
  group_by(region, country) %>%
  mutate(
    across(
      c(internet, fixed, mobile),
      list(`1` = ~ lag(., 1)),
      .names = "{str_sub(.col, 1, 1)}{.fn}"
    )
  ) %>% 
  ungroup()

# Centering predictors within and between countries
itu <- itu %>%
  group_by(country) %>%
  mutate(
    i1_cb = mean(i1, na.rm = TRUE), # Country average
    i1_cw = i1 - i1_cb, # Year-specific deviation from country average
    f1_cb = mean(f1, na.rm = TRUE),
    f1_cw = f1 - f1_cb,
    m1_cb = mean(m1, na.rm = TRUE),
    m1_cw = m1 - m1_cb,
  ) %>%
  # Grand-mean center the country averages
  ungroup() %>% 
  mutate(
    across(c(i1_cb, f1_cb, m1_cb), ~ . - mean(., na.rm = T))
  )

# Transform internet edge cases
itu <- itu %>%
  mutate(internet = if_else(internet == 0, .0000001, internet)) %>%
  mutate(internet = if_else(internet == 1, .9999999, internet))

# Stack outcome data
dat <- bind_rows(gwp, gbd)

# Merge predictors to outcomes
dat <- left_join(dat, itu)

# Center and scale year
dat <- dat %>%
  mutate(year = (year - 2010) / 10)

# Identify values to use in mv models
# Indicate unique ITU values for MV model
dat <- dat %>%
  # Unique values exist for each country-year, but must be given for each
  # outcome
  group_by(region, country, year, outcome) %>%
  mutate(itu = 1:n() == 1) %>%
  ungroup()

# Ensure that factors are contrast coded
dat <- dat %>%
  mutate(across(c(sex, age), factor))
options(contrasts = c(unordered = "contr.sum", ordered = "contr.poly"))

# Model outcomes on the percentage (0-100) scale. Note internet is still 0-1
dat <- dat %>%
  mutate(across(c(val, se), ~ . * 100))

# Arrange on outcome
oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)
dat <- dat %>%
  mutate(outcome = factor(outcome, levels = oo)) %>% 
  arrange(dat, outcome, region, country, year)

dat %>%
  group_by(outcome) %>%
  summarise(
    min = min(val, na.rm = TRUE),
    max = max(val, na.rm = TRUE),
    internet_min = min(internet, na.rm = TRUE),
    internet_max = max(internet, na.rm = TRUE)
  ) %>%
  kable(digits = 7) %>% 
  kable_custom()
```

We then create the lagged and centered GWP internet predictor. Note that this is at the level of demographics, not just a country average like in ITU.

```{r}
# Take only relevant variables to make this easier
gwpi <- dat %>% 
  filter(outcome == "Life_satisfaction") %>% 
  filter(year >= .5) %>% 
  select(region, country, year, sex, age, year, gwpi)

# Create lagged GWP internet
gwpi <- gwpi %>% 
  arrange(region, country, sex, age, year) %>% 
  group_by(region, country, sex, age) %>% 
  mutate(g1 = lag(gwpi, 1))

# Center GWP internet
gwpi <- gwpi %>%
  group_by(country) %>%
  mutate(
    g1_cb = mean(g1, na.rm = TRUE), # Country average
    g1_cw = g1 - g1_cb, # Year-specific deviation from country average
  ) %>%
  # Grand-mean center the country averages
  ungroup() %>% 
  mutate(
    g1_cb = g1_cb - mean(g1_cb, na.rm = TRUE)
  )

# Join to data (all outcomes)
dat <- left_join(dat, select(gwpi, -gwpi))
```

@@

```{r model1-itu-bf}
#| echo: true
# Define a model of internet proportion over year
bf_itu <- bf(
  internet | resp_subset(itu) ~
  year +
    (year |c| country) +
    (year |r| region),
  family = Beta()
)
```

@@

```{r model1-val-bf}
#| echo: true
# Meta-regression model of outcomes
bf_1 <- bf(
  val | resp_se(se, sigma = TRUE) ~
  year * sex * age +
    (year * sex * age | c | country) +
    (year * sex * age | r | region),
  family = gaussian()
)
```

@@

```{r model1-selfharm-se}
#| echo: true
dat <- dat %>% 
  mutate(se = if_else(outcome == "Selfharm", 0, se))
```

@@

We then create a table where each row is a dataset and a model to be fitted, and then fit the models row-by-row. This takes about a day for the six outcomes.

```{r fit-model-1}
#| cache: false
fits <- dat %>%
  # Could nest but need this to include group key in nested tibble
  group_by(outcome) %>%
  summarise(data = list(cur_data_all()))

fits <- fits %>%
  mutate(
    fit1 = map2(
      data, outcome,
      ~brm(
        formula = bf_itu + bf_1 + set_rescor(FALSE),
        data = .x,
        chains = hmc$chains,
        cores = hmc$cores,
        threads = hmc$threads,
        iter = hmc$iter,
        warmup = hmc$warmup,
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta,
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("models/brm-{.y}-1-internet")
      )
    )
  )
```

@@

```{r fit-model-2}
#| cache: false
#| results: hide
# Model formula is only a small update to Model 1
bf_2 <- bf(
  val | resp_se(se, sigma = TRUE) ~
  year * sex * age + i1_cw * sex * age + i1_cb +
    (year * sex * age + i1_cw * sex * age | country) +
    (year * sex * age + i1_cw * sex * age | region),
  family = gaussian()
)

# As above with model 1
# Note we don't need to save these in the R environment, just files
fits %>%
  mutate(
    fit2 = map2(
      data, outcome,
      ~brm(
        formula = bf_2,
        data = .x,
        chains = hmc$chains,
        cores = hmc$cores,
        threads = hmc$threads,
        iter = hmc$iter,
        warmup = hmc$warmup,
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta,
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("models/brm-{.y}-2-internet")
      )
    )
  )
```

@@

```{r fixed-mobile}
#| cache: false
#| results: hide
bf_2_f <- bf(
  val | resp_se(se, sigma = TRUE) ~
  year * sex * age + i1_cw * sex * age + f1_cb +
    (year * sex * age + i1_cw * sex * age | country) +
    (year * sex * age + i1_cw * sex * age | region),
  family = gaussian(),
  unused = ~outcome
)

bf_2_m <- bf(
  val | resp_se(se, sigma = TRUE) ~
  year * sex * age + i1_cw * sex * age + m1_cb +
    (year * sex * age + i1_cw * sex * age | country) +
    (year * sex * age + i1_cw * sex * age | region),
  family = gaussian(),
  unused = ~outcome
)

bf_2_g <- bf(
  val | resp_se(se, sigma = TRUE) ~
  year * sex * age + i1_cw * sex * age + g1_cb +
    (year * sex * age + i1_cw * sex * age | country) +
    (year * sex * age + i1_cw * sex * age | region),
  family = gaussian(),
  unused = ~outcome
)

fits %>%
  mutate(
    fit2_f = map2(
      data, outcome,
      ~brm(
        formula = bf_2_f,
        data = .x,
        chains = hmc$chains,
        cores = hmc$cores,
        threads = hmc$threads,
        iter = hmc$iter,
        warmup = hmc$warmup,
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta,
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("models/brm-{.y}-2-fixed")
      )
    )
  )

fits %>%
  mutate(
    fit2_m = map2(
      data, outcome,
      ~brm(
        formula = bf_2_m,
        data = .x,
        chains = hmc$chains,
        cores = hmc$cores,
        threads = hmc$threads,
        iter = hmc$iter,
        warmup = hmc$warmup,
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta,
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("models/brm-{.y}-2-mobile")
      )
    )
  )

fits %>%
  mutate(
    fit2_g = map2(
      data, outcome,
      ~brm(
        formula = bf_2_g,
        data = .x,
        chains = hmc$chains,
        cores = hmc$cores,
        threads = hmc$threads,
        iter = hmc$iter,
        warmup = hmc$warmup,
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta,
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("models/brm-{.y}-2-gwpinternet")
      )
    )
  )
```

```{r}
fits$data[[1]] %>% select(country, year, g1_cb) %>% drop_na(g1_cb)
```

>>>>

# Models

```{r setup}
#| cache: false
# Packages
library(scales)
library(ggtext)
library(lubridate)
library(kableExtra)
library(ggstance)
library(mgcv)
library(cmdstanr)
library(ellipse)
library(ggh4x)
library(tidybayes)
library(ggdist)
library(posterior)
library(parameters)
library(emmeans)
library(broom)
library(brms)
library(tidyverse)
```

@@

@@

```{r model1-itu-bf}
#| echo: true
```

@@

```{r model1-val-bf}
#| echo: true
```

@@

@@

```{r fit-model-1}
#| cache: false

```

@@

```{r fit-model-2}
#| cache: false
#| results: hide

```

@@

```{r fixed-mobile}
#| cache: false
#| results: hide

```
